Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-05-27 compare.tex
test for compare

% cpedoc.tex V2.0, 13 May 2010

\documentclass[times]{cpeauth}

\usepackage{moreverb}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2012}
%\documentclass[acmjacm,acmnow]{acmtrans2m}
%\documentclass[conference,final]{IEEEtran}
%\documentclass[10pt,letterpaper]{article}
%\documentclass[preprint,12pt]{article}
%\documentclass[11pt,letterpaper]{article}

\usepackage{graphicx}
\usepackage{url}
\usepackage{color}
\usepackage{enumerate}
\usepackage{longtable}
%\usepackage{authblk} % Is used internally within the CPE Class
%\usepackage[firstpage]{draftwatermark}

%\SetWatermarkLightness{0.7}
%\SetWatermarkScale{4}

%\usepackage{pdfsync}
% \acmVolume{}
% \acmNumber{}
% \acmYear{}
% \acmMonth{}

%\usepackage{sectsty}
%\usepackage{setspace}

% \usepackage{ifpdf}
% \usepackage{wrapfig}
% %\usepackage[cm]{fullpage}
% \usepackage{textcomp}
% \usepackage{srcltx}
% \usepackage{fancyhdr}
% \usepackage{setspace}
\usepackage{lscape}
% \usepackage{longtable}
% \usepackage{paralist}

%\usepackage{hyperref}
\usepackage[novbox]{pdfsync}

% \usepackage{ams}

% \ifpdf
%              \usepackage[pdftex,
%               colorlinks=true,
%               linkcolor=red,
%               citecolor=red,
%               filecolor=red,
%               ]{hyperref}
% \else
%   \usepackage[hypertex]{hyperref}
% \fi

% Space saving
%\usepackage[small,compact]{titlesec}
%\usepackage[small,it]{caption}

% \renewcommand\floatpagefraction{.9}
% \renewcommand\topfraction{.9}
% \renewcommand\bottomfraction{.9}
%\renewcommand\textfrare{.1}

\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

%\textwidth = 6in
%\textheight = 8.5 in
%\oddsidemargin = 0.0 in
%\evensidemargin = 0.0 in

% \topmargin = 0.0 in
% \headheight = 0.0 in
% \headsep = 0.0 in
%\parskip = 0.0in
%\parindent = 0.5cm
\parskip = 0.04in
% \parindent = 0.0cm
% \textfloatsep = 0.1in

% \usepackage[draft,pdftex,colorlinks=true, linkcolor=blue,citecolor=blue,
%        urlcolor=blue]{hyperref}

% \ifpdf
% \pdfinfo{ /Author (Shantenu Jha et al.)  /Title (Programming Abstractions for Large-Scale Distributed Applications) }
% \fi

%\newcommand{\projectnamefull}{\textit{Programming Abstractions for
%    Large-Scale Distributed Applications} }

\newcommand{\upup}{\vspace*{-0.5em}}
\newcommand{\up}{\vspace*{-0.25em}}
\newcommand{\I}[1]{\textit{#1}}
\newcommand{\B}[1]{\textbf{#1}}
\newcommand{\T}[1]{\texttt{#1}}
\newcommand{\BI}[1]{\B{\I{#1}}}

 \newenvironment{shortlist}{
        \vspace*{-0.8em}
  \begin{itemize}
  \setlength{\itemsep}{-0.3em}
}{
  \end{itemize}
        \vspace*{-0.6em}
}

 \newenvironment{shortenum}{
        \vspace*{-0.8em}
  \begin{enumerate}
  \setlength{\itemsep}{-0.3em}
}{
  \end{enumerate}
        \vspace*{-0.6em}
}

% \ifpdf
%   \DeclareGraphicsExtensions{.pdf, .png, .jpg}
% \else
%   \DeclareGraphicsExtensions{.ps, .eps}
% \fi

\definecolor{orange}{rgb}{1.0,0.3,0.0}
\definecolor{violet}{rgb}{0.75,0,1}
\definecolor{darkgreen}{rgb}{0,0.6,0}
\definecolor{cyan}{rgb}{0.2,0.7,0.7}
\definecolor{blueish}{rgb}{0.2,0.2,0.8}

\long\def\comment#1{{\bf \textcolor{magenta}{\bf #1}}}
\long\def\ccomment#1{{\bf \textcolor{blue}{\bf #1}}}
\newcommand{\C}{\comment}
\newcommand{\CC}{\ccomment}

\newcommand{\yes}{$\bullet$}

\long\def\finalcomment#1{{\bf \textcolor{red}{\bf #1}}}
\newcommand{\CF}{\finalcomment}

\newif\ifdraft
\drafttrue
\ifdraft
 \newcommand{\katznote}[1]{ {\textcolor{magenta}    { ***Dan:      #1 }}}
 \newcommand{\rananote}[1]{ {\textcolor{cyan}    { ***Omer:     #1 }}}
 \newcommand{\omernote}[1]{ {\textcolor{cyan}    { ***Omer:     #1 }}}
 \newcommand{\note}[1]{ {\textcolor{red}    { #1 }}}
 \newcommand{\jhanote}[1]{  {\textcolor{blue}     { ***Shantenu: #1 }}}
 \newcommand{\jonnote}[1]{  {\textcolor{red}      { ***Jon: #1 }}}
 \newcommand{\alnote}[1]{  {\textcolor{darkgreen}      { ***Andre: #1 }}}
 \newcommand{\nchnote}[1]{  {\textcolor{orange}      { ***Neil: #1 }}}
 \newcommand{\ysnote}[1]{  {\textcolor{violet}      { ***Yogesh: #1 }}}
 \newcommand{\sdnote}[1]{  {\textcolor{blueish}      { ***Simon: #1 }}}
% \usepackage[pdftex,colorlinks=true, linkcolor=blue,citecolor=blue,
%       urlcolor=blue]{hyperref}
\else
 \newcommand{\katznote}[1]{}
 \newcommand{\rananote}[1]{}
 \newcommand{\note}[1]{}
 \newcommand{\jhanote}[1]{}
 \newcommand{\jonnote}[1]{}
\newcommand{\alnote}[1]{}
 \newcommand{\nchnote}[1]{}
 \newcommand{\ysnote}[1]{}
\fi

@@

\address{\affilnum{1}Rutgers University\break
\affilnum{2}University of Edinburgh\break
\affilnum{3}University of St.~Andrews\break
\affilnum{4}University of Chicago\break
\affilnum{5}Argonne National Laboratory\break
\affilnum{6}Cardiff University\break
\affilnum{7}Indian Institute of Science\break
}

@@

Pan-STARRS, a contemporary sky survey with similar goals as LSST,  uses a scripted programming model
on a cluster for image processing of telescope frames to extract object attributes as a CSV
file.
%\alnote{Is the workflow part still part of the app as defined before with the stages: data transformation \& decision and storage?} \katznote{I think this is data transformation.}\alnote{ok}
This is followed by followed by scientific workflows, based on Trident, that load the files
into individual databases daily, and merges them with a master distributed-database on a weekly basis on an
HPC cluster~\cite{Simmhan:Building:2009}. Data extraction and mining operations by astronomers are composed using a pseudo-SQL
language with user defined functions and submitted for execution on the master database using a batch system.

@@

The application~\cite{modisAzure, modis:ipdps:2010} is used to handle the processes of downloading data from different
sites and integrating them. The integration consists of reprojecting the data from one map projection to another,
spatial and temporal resampling, and gap filling. The application itself is a loosely-coupled data
pipeline that coordinates these processes.

@@

In this section, we focus on the {\em software} infrastructure that is
provided to support D3 applications, specifically components that deal
with distributed or dynamic data. The components that make up the infrastructure can be categorized
based on the function provided, namely (i) infrastructure that supports {\it data management},
(ii) infrastructure that supports {\it data analysis}, and (iii) infrastructure that supports the \emph{coordination within applications}.
Each of these components may be present at three different levels of the infrastructure: programming frameworks, services and tools (e.g., MapReduce, workflow engines); middleware and platform services (e.g., databases, message queues); and system-level
software capabilities (e.g., notifications, file system consistency).

@@

{\em Data Sources} are used in the acquisition stage as inputs of data
into the application. Broadly speaking these sources include types such as
files, servers, sensors, instruments, simulations and models.
Additionally, databases are often considered as an intermediate data
source in some systems. The dynamic characteristics are primarily
due to availability. These are often accessed in the Data Generation stage of the application
(\figurename~\ref{fig:figures_application-stages}).

{\em Data Storage} components are used to provide persistent access to
data by the application. They may be structured (e.g., databases,
No-SQL tables) or unstructured (e.g., collections of audio
recordings). The dynamic characteristics are primarily due to the
level of persistence, e.g., temporary, durable, permanent (archived or
published), or replicated; and also in the structure of the data
stored, e.g., schema free tables as opposed to relational tables. There
may also be dynamic characteristics of storage due to rate of growth
of data, e.g., a field survey may bring in a lot of data during summer
but none during other times. Some storage components like files and databases may also serve as data
sources. These are commonly used in the Data Storage stage of the application (\figurename~\ref{fig:figures_application-stages}),
but may also be present at the boundaries between the stages for transient staging of data.

@@

{\em Data Movement} components are used for bulk or continuous
movement of data between parts of the application, or to and from the
application, % (often called staging) Data movement may be explicit
and visible to the application (e.g., staging), or implicit (e.g.,
automatic replication of files). Components may operate on streams or
blocks (files). Dynamic characteristics include partitioning,
scheduling and planning (e.g., edge caching of ``hot'' data, where
datasets that are being used frequently during some period are
replicated close to the application). Data movement can be considered
to be real-time (i.e., as it happens, e.g., sensor data), `as real-time'
(e.g., a storage buffer has been introduced into the transport system,
but the data source is still real-time, as with time-series data), or
asynchronous (where the ordering of the data movement is independent
of the data itself).  Latency may be considered as an important
characteristic of the data movement infrastructure depending on the
application, e.g., real-time processing of sensor data for disaster
analysis. Other dynamic aspects of data movement include data rates,
their variability from data sources, and the ``freshness'' of
data. These infrastructure appear on the edges between the application
stages.

{\em Data Discovery} components assist the access to and use of data
sources, and include Catalogs, Information Services, and Metadata
stores. The dynamic characteristics pertain to the evolution or
recalibration of information held in these components. These components are orthogonal to the
applications stages and can be used by any of the
stages to guide their selection or identification of incoming data sources and outgoing data sinks
or storage.
%\ysnote{rate of growth of data may be an issue.
%Catalogs that perform nightly ``crawling'' or data loading (e.g., sky surveys) may not be suitable if the
%latest data needs to be accessible to applications. Catalog update will have to keep up with data
%update. However, batch processing applications can survive less frequent updates.}

@@

Analysis components include functionality such as comparison,
statistical analysis, event analysis (i.e., analyzing information
derived from data rather than raw data) and visualization. These components leverage
the computational aspects of the infrastructure.
Based on the application stages outlined in
\figurename~\ref{fig:figures_application-stages}, we identify a number
of infrastructure services for data analysis that can be associated with these stages.

@@

%\ysnote{addresses Andre's commment}
{\em Analytics and visualization} components help extract domain knowledge and assist with
exploration of the data. These are often closely supervised by the end-user and may even be
interactive. The proliferation of ``big data'' has caused data mining libraries, machine-learning
toolboxes and visualization environments to become part of the infrastructure, some of which are
deployed for specific domains but others which are more generic (e.g., Weka, GraphLab, Apache Mahout). These components are associated
with the Data Processing stage.

@@

{\em Orchestration} components handle the management and
choreography of data placement, code placement, communication, and
messaging to allow complex analyses to be performed on distributed
data sources. These are often composed as workflows or Directed Acyclic Graphs (DAGs) that capture
the execution dependency between
components. They may also be dynamic if they have the capability to
adapt their behavior based on earlier parts of the workflow or can be
steered based on external notification received during execution. In
addition to supporting choreography, while often making use of one or more
resource manager/scheduler components, an orchestration component may also
include specialist support for conversion and analysis components
discussed above.

{\em Bulk Data-parallel} frameworks use ``Bulk'' operators to express computations over 
collections of data which are executed in a loosely-coupled, data-parallel manner.
The {\em MapReduce} paradigm~\cite{MapReduce} and its variants~\cite{twister,hmr} are the most
prominent among these, with recent ones like Pregel and Apace Giraph~\cite{pregel} supporting specific domains like
graph analytics. MapReduce forces programmers to model computation in two stages: 
a \emph{map} phase in which a transformation is applied to every element of a collection independently of every other;  and a \emph{reduce} phase in which the transformed results 
are aggregated through a binary function that is associative and commutative. 
These properties serve to guarantee that the MapReduce computation can be efficiently 
parallelized, by performing the map phase concurrently and
then reducing without the reduction order affecting the final result.

@@

\hline	
\multicolumn{2}{|l|}{{\textbf{Application Name}}} \\ \hline	
%\nchnote{For example, workload decomposition and distribution must be determined dynamically in order to optimally use resources, e.g., compute resource selection can be based on data location, processing profile and/or network capacity. Workflows? Resource selection?}
Infrastructure to support Dynamic Properties & Infrastructure to support Distributed Properties
 \\ \hline
%\begin{footnotesize}
%	 \begin{longtable}[h!]{|p{6.7cm}|p{6.7cm}|}
%		\caption{Infrastructure used to support dynamic and distributed properties of application scenarios} \label{tab:inf_app_mapping} \\
%		\hline
%		\textbf{Infrastructure to support Dynamism}
%                &\textbf{Infrastructure to support Distributedness} \\		
%		 \endhead
		\hline	
\multicolumn{2}{|l|}{{\textbf{Next Generation Sequencing (NGS) Analytics} (\S\ref{bioSilvia})}} \\ \hline	
%\nchnote{For example, workload decomposition and distribution must be determined dynamically in order to optimally use resources, e.g., compute resource selection can be based on data location, processing profile and/or network capacity. Workflows? Resource selection?}
Dynamic workload decomposition and distribution for resource optimization \emph{(Orchestration)}
		&
Distributing data from source (generation) to the computing destination (analysis) \emph{(Data
  Discovery, Movement)}
% 		No unifying single requirement or single solution at present (or possible at all?)
% 		Normally data local and CI doesn't easily support movement of data to compute
% \ysnote{We may want to remove the preceding lines since we're talking about existing infrastructure, not missing ones at this point.}
% naturally distributed: data source (NGS machine storage) and compute on
% 		different resources (typically not co-located); loosely coupled (many
%		machines, different repositories)
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{ATLAS (an example of use of the WLCG) Reconstruction app} (\S\ref{WLCGSteve})}} \\ \hline
Manage gradual growth of data using SRM \emph{(Data Storage, Access)}
% Management, subset replication and partitioning to go inf components: data replication, (gLite), data movement
% specify SRM as standard and then they leave choice of implementation to the organisations as part of consortium.
% New raw data is collected
% 	    steadily (2) \note{timescale needs to worked out. What happens if something is found?}
&
Management, subset replication, data movement and partitioning using gLite and SRM \emph{(Data
  Storage, Access, Movement, Discovery)}
% , data movement
% natural, (sub-set) replicated, loosely coupled \note{Is natural the
% 	   right word? Artificial
% 		distribution (the source is natural, but the intermediate points are
% 		artificial), data not partitioned}\ysnote{These details are not fully present in the actual application description of its dynamism (Sec 2.2). Should we trim there and move some content there? }
              \\
		\hline
\multicolumn{2}{|l|}{{\textbf{LSST} (\S\ref{astroAdam})}} \\ \hline

Steering observations, Databases to manage data growth \emph{(Notification, Data Storage)}
% DE decisions based on observed data (2, 5) \ysnote{what do these numbers mean? specific
             %  characteristics?}
            &
Coordinate classification, distributed databases, data movement to distributed observatories
\emph{(Notification, Data Access, Data Movement, Data Storage)} %Analysis, % , natural distribution of telescopes
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{SOA Astronomy Applications} (\S\ref{astroSOAAdam})}} \\ \hline

Catalog contents updated, Query mix changes \emph{(Data Discovery)}
% data is continuously updated, thus dynamic data (pre-defined schedules or on-demand), queries change (7)\note{timescale clarification!}
             &
Distributed VO Services for data storage and access \emph{(Data Discovery, Data Access, Data Movement)}
% moving data sources, services and storage are distributed across institutions
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{Cosmic Microwave Background} (\S\ref{astroJulian})}} \\ \hline	
	      Dynamic resource selection  \emph{(Data Discovery)}
%dynamic utilization of resources (4)
& Compute on distributed HPC centers \emph{(Data Movement)}
% (sub-set) replication (during application run) \note{What phases of the application do we consider? data acquisition? app run?}\ysnote{Can any of the supercomputers be selected? If so, the data is moved to distributed compute (or it is dynamic resource selection)}
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{Sensor Network Application} (\S\ref{sensorSimon})}} \\ \hline	
	     Notification of sensor data when in range, data rates and volumes \emph{(Data Movement, Notification)} %data rate, volume dynamic (1-2)
&Aggregating distributed sensor data \emph{(Data Movement, Notification)}  \\
		\hline
\multicolumn{2}{|l|}{{\textbf{Climate (ESGF)} (\S\ref{climateDan})}} \\ \hline	
	      Manage data growth % New data generated over time (ESGF-wide - is this in scope of the application?)
             &
Distributed, replicated catalogs using ESGF software stack (GridFTP, wget/HTTP) \emph{(Data Access,
  Data Discovery, Data Movement, Data Conversion)} \\
  % multiple climate centers (natural), users can request data from ESGF (sub-set replication), coupled
% Can come in via website - cookie / OpenID
% Come in through command line script - grid cert
		\hline
\multicolumn{2}{|l|}{{\textbf{Interactive Exploration of Environmental Data} (\S\ref{envJon})}} \\ \hline	
On-demand data subsetting, caching and versioning, real-time instrument data \emph{(Data Discovery, Data
  Movement, Data Sources)}
% at moment services are hosted on a single address. There may be farming for load balancing in the future.
% data rates change (1),
% 	    data inputs/outputs change (3), queries change (7)
		&
Distributed Data Services, Access Restrictions \emph{(Data
  Movement, Data Access)}

@@

\\
		\hline
	    % Dynamic Environmental Data on the Web &data operated on changes
	    %     (realtime feeds) (2)
	    %     &natural (sensor sources, simulations), (sub-set) replication\\
	    %     \hline
\multicolumn{2}{|l|}{{\textbf{PowerGrids} (\S\ref{power})}} \\ \hline

@@

		\hline
\multicolumn{2}{|l|}{{\textbf{Fusion (ITER)} (\S\ref{fusionScott})}} \\ \hline	
	     Stream and transform data between simulation sites on-demand \emph{(Data Movement, Enrichment)}
%  (multiple distinct data
	    % flows), on-the-fly
	    % transformation (8)
	    &
Run analysis on available, distributed resources \emph{(Data Movement)} %Analysis
% natural, coupled, transient sub-set replicated
            \\
		\hline
\multicolumn{2}{|l|}{{\textbf{Industrial Incident Notification and Response} (\S\ref{Kees})}} \\ \hline	
	     Acquire different data sources and rates, Processing determined by data
             \emph{(Data Sources, Data Access, Notification, Orchestration)}   %Analysis,  % (1),
	    % data itself changes (2), dynamic data sources (3), processing changes
	    % based on data
            %(5)
&
Distributed data sources, Analysis by agents across institutions, Access control \emph{(Data Sources, Data
  Access)} % , Analysis
% natural, coupled
            \\
		\hline
\multicolumn{2}{|l|}{{\textbf{MODIS Data Processing} (\S\ref{modisKeith})}} \\ \hline

@@

\multicolumn{2}{|l|}{{\textbf{Distributed Network Intrusion Detection} (\S\ref{detectionJon})}} \\ \hline	
Analyze dynamic network traffic rates \emph{(Data Movement)} % , Analysis
	    %  data rate variable (1), data
	    % highly dynamic changing (2), application responds to data (5),
	    % on-the-flight processing (8)
            &
Run analysis at different sites, Distribute generated data \emph{(Data Movement)} % Analysis,
% natural, data/compute tightly coupled
\\
		\hline
%	 \end{longtable}
%\end{footnotesize}

@@

Revisiting the overall issue of infrastructure, which we defined as
the hardware and software needed to run the applications, we
note that the specific hardware is not a primary issue.  The
applications we have discussed all use a mix of the systems that exist
today, which have aspects of web, grid, and cloud, and can be
considered a hybrid of the three.  The key infrastructure components
are really the software and services on top of the hardware.  In this
section, we have discussed infrastructure to support data management and
data analysis, as well as the tools and systems used to program
applications, and the specific components used for the application in
\S\ref{sec:scenarios}.  However, there is not much commonality in the
software and services that the applications use, so the software and programming infrastructure
we have described here is much more broad than what is used by the
individual applications we have studied.

@@

In this section we attempt to summarize four key themes from previous
sections: (i) the source and nature of dynamic data that we observe in
the application surveyed in \S\ref{sec:scenarios}; (ii) the types of
data distribution that we observe in such applications; (iii) the
general characteristics that can be associated with D3 science
applications; (iv) the characteristics of a computational
infrastructure that may be most suited for D3 science
applications. Whereas (i) and (ii) are primarily observations that we
can make on existing applications that we have discussed in
\S\ref{sec:scenarios}, (iii) and (iv) attempt to synthesize general
trends that can be applied to other applications we have not
considered. \katznote{should this be split into two subsections then? Perhaps Observations \& Trends?}

%\katznote{Some text is needed here to introduce the discussion subsection.}
%\rananote{Added some text above -- does this make sense Dan?}

>>>>

\include{header}

@@

\address{\affilnum{1}Rutgers University\break
\affilnum{2}University of Edinburgh\break
\affilnum{3}University of St.~Andrews\break
\affilnum{4}University of Chicago\break
\affilnum{5}Argonne National Laboratory\break
\affilnum{6}Cardiff University\break
\affilnum{7}University of Southern California\break
}

@@

Pan-STARRS, a contemporary sky survey with similar goals as LSST,  uses a scripted programming model
on a cluster for image processing of telescope frames to extract object attributes as a CSV
file.
%\alnote{Is the workflow part still part of the app as defined before with the stages: data transformation \& decision and storage?} \katznote{I think this is data transformation.}\alnote{ok}
This is followed by followed by scientific workflows, based on Trident, that load the files
into individual databases daily, and merges them with a master distributed-database on a weekly basis on an
HPC cluster. Data extraction and mining operations by astronomers are composed using a pseudo-SQL
language with user defined functions and submitted for execution on the master database using a batch system.

@@

The application~\cite{modisAzure} is used to handle the processes of downloading data from different
sites and integrating them. The integration consists of reprojecting the data from one map projection to another,
spatial and temporal resampling, and gap filling. The application itself is a loosely-coupled data
pipeline that coordinates these processes.

@@

In this section, we focus on the {\em software} infrastructure that is
provided to support D3 applications, specifically components that deal
with distributed or dynamic data. The components that make up the infrastructure can be categorized
based on the function provided, namely (i) infrastructure that supports {\it data management},
(ii) infrastructure that supports {\it data analysis}, and (iii) infrastructure that supports the structure of the applications.
Each of these components may be present at three different levels of the infrastructure: programming frameworks, services and tools (e.g., MapReduce, workflow engines); middleware and platform services (e.g., databases, message queues); and system-level
software capabilities (e.g., notifications, file system consistency).

@@

{\em Data Sources} are used in the acquisition stage as inputs of data
into the application. Broadly speaking they include types such as
files, servers, sensors, instruments, simulations and models.
Additionally, databases are often considered as an intermediate data
source in types of system. The dynamic characteristics are primarily
due to availability. These are often accessed in the Data Generation stage of the application
(\figurename~\ref{fig:figures_application-stages}).

{\em Data Storage} components are used to provide persistent access to
data by the application. They may be structured (e.g., databases,
No-SQL tables) or unstructured (e.g., collections of audio
recordings). The dynamic characteristics are primarily due to the
level of persistence, e.g., temporary, durable, permanent (archived or
published), or replicated; and also in the structure of the data
stored, e.g., schema free tables as opposed to relational tables. There
may also be dynamic characteristics of storage due to rate of growth
of data, e.g., a field survey may bring in a lot of data during summer
but none during other times. These are commonly used in the Data Storage stage of the application (\figurename~\ref{fig:figures_application-stages}),
but may also be present at the boundaries between the stages for transient staging of data.

@@

{\em Data Movement} components are used for bulk or continuous
movement of data between parts of the application, or to and from the
application. % (often called staging) Data movement may be explicit
and visible to the application (e.g., staging), or implicit (e.g.,
automatic replication of files). Components may operate on streams or
blocks (files). Dynamic characteristics include partitioning,
scheduling and planning (e.g., edge caching of ``hot'' data, where
datasets that are being used frequently during some period are
replicated close to the application). Data movement can be considered
to be real-time (i.e., as it happens, e.g., sensor data), as real-time
(e.g., a storage buffer has been introduced into the transport system,
but the data source is still real-time, as with time-series data), or
asynchronous (where the ordering of the data movement is independent
of the data itself).  Latency may be considered as an important
characteristic of the data movement infrastructure depending on the
application e.g. real-time processing of sensor data for disaster
analysis. Other dynamic aspects of data movement include data rates,
their variability from data sources, and the ``freshness'' of
data. These infrastructure appear on the edges between the application
stages.

{\em Data Discovery} components assist the access to and use of data
sources, and include Catalogs, Information Services, and Metadata
stores. The dynamic characteristics pertain to the evolution or
recalibration of information held in these components. These components are orthogonal to the
applications stages and can be used by any of the
stages to guide their selection or identification of incoming data sources and outgoing data sinks.
%\ysnote{rate of growth of data may be an issue.
%Catalogs that perform nightly ``crawling'' or data loading (e.g., sky surveys) may not be suitable if the
%latest data needs to be accessible to applications. Catalog update will have to keep up with data
%update. However, batch processing applications can survive less frequent updates.}

@@

Analysis components include functionality such as comparison,
statistical analysis, event analysis (i.e., analyzing information
derived from data rather than raw data) and visualization. These components leverage
the computational aspects of the infrastructure.
Based on the application stages outlined in
\figurename~\ref{fig:figures_application-stages}, we identify a number
of infrastructure services that can be associated with these stages.

@@

%\ysnote{addresses Andre's commment}
{\em Analytics and visualization} components help extract domain knowledge and assist with
exploration of the data. These are often closely supervised by the end-user and may even be
interactive. The proliferation of ``big data'' has caused data mining libraries, machine-learning
toolboxes and visualization environments to become part of the infrastructure, some of which are
deployed for specific domains but others which are more generic (e.g., Weka, GraphLab). These components are associated
with the Data Processing stage.

@@

{\em Orchestration} components handle the management and
choreography of data placement, code placement, communication, and
messaging to allow complex analyses to be performed on distributed
data sources. They may also be dynamic if they have the capability to
adapt their behavior based on earlier parts of the workflow or can be
steered based on external notification received during execution. In
addition to supporting choreography, often making use of one or more
resource manager/scheduler components, an orchestration component may also
include specialist support for conversion and analysis components
discussed above.

{\em Bulk Data-parallel} frameworks use ``Bulk'' operators to express computations over 
collections of data which are executed in a loosely-coupled, data-parallel manner.
The {\em MapReduce} paradigm~\cite{MapReduce} and its variants~\cite{twister,hmr} are the most prominent among these. MapReduce forces programmers to model computation in two stages: 
a \emph{map} phase in which a transformation is applied to every element of a collection independently of every other;  and a \emph{reduce} phase in which the transformed results 
are aggregated through a binary function that is associative and commutative. 
These properties serve to guarantee that the MapReduce computation can be efficiently 
parallelized, by performing the map phase concurrently and
then reducing without the reduction order affecting the final result.

@@

\hline	
\multicolumn{2}{|l|}{{\textbf{Application Name}}} \\ \hline	
%\nchnote{For example, workload decomposition and distribution must be determined dynamically in order to optimally use resources, e.g., compute resource selection can be based on data location, processing profile and/or network capacity. Workflows? Resource selection?}
Infrastructure for Dynamic Properties & Infrastructure for Distributed Properties
 \\ \hline
%\begin{footnotesize}
%	 \begin{longtable}[h!]{|p{6.7cm}|p{6.7cm}|}
%		\caption{Infrastructure used to support dynamic and distributed properties of application scenarios} \label{tab:inf_app_mapping} \\
%		\hline
%		\textbf{Infrastructure to support Dynamism}
%                &\textbf{Infrastructure to support Distributedness} \\		
%		 \endhead
		\hline	
\multicolumn{2}{|l|}{{\textbf{Next Generation Sequencing (NGS) Analytics}}} \\ \hline	
%\nchnote{For example, workload decomposition and distribution must be determined dynamically in order to optimally use resources, e.g., compute resource selection can be based on data location, processing profile and/or network capacity. Workflows? Resource selection?}
Dynamic workload decomposition and distribution for resource optimization \emph{(Orchestration)}
		&
Distributing data from source (generation) to the computing destination (analysis) \emph{(Data
  Discovery, Movement)}
% 		No unifying single requirement or single solution at present (or possible at all?)
% 		Normally data local and CI doesn't easily support movement of data to compute
% \ysnote{We may want to remove the preceding lines since we're talking about existing infrastructure, not missing ones at this point.}
% naturally distributed: data source (NGS machine storage) and compute on
% 		different resources (typically not co-located); loosely coupled (many
%		machines, different repositories)
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{ATLAS (an example of use of the WLCG) Reconstruction app}}} \\ \hline
Manage gradual growth of data using SRM \emph{(Data Storage, Access)}
% Management, subset replication and partitioning to go inf components: data replication, (gLite), data movement
% specify SRM as standard and then they leave choice of implementation to the organisations as part of consortium.
% New raw data is collected
% 	    steadily (2) \note{timescale needs to worked out. What happens if something is found?}
&
Management, subset replication, data movement and partitioning using gLite and SRM \emph{(Data
  Storage, Access, Movement, Discovery)}
% , data movement
% natural, (sub-set) replicated, loosely coupled \note{Is natural the
% 	   right word? Artificial
% 		distribution (the source is natural, but the intermediate points are
% 		artificial), data not partitioned}\ysnote{These details are not fully present in the actual application description of its dynamism (Sec 2.2). Should we trim there and move some content there? }
              \\
		\hline
\multicolumn{2}{|l|}{{\textbf{LSST}}} \\ \hline

Steering observations, Databases to manage data growth \emph{(Notification, Data Storage)}
% DE decisions based on observed data (2, 5) \ysnote{what do these numbers mean? specific
             %  characteristics?}
            &
Coordinate classification, distributed databases, data movement to distributed observatories
\emph{(Notification, Data Access, Data Movement, Data Storage)} %Analysis, % , natural distribution of telescopes
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{SOA Astronomy Applications}}} \\ \hline

Catalog contents updated, Query mix changes \emph{(Data Discovery)}
% data is continuously updated, thus dynamic data (pre-defined schedules or on-demand), queries change (7)\note{timescale clarification!}
             &
Distributed VO Services for data storage and access \emph{(Data Discovery, Data Access, Data Movement)}
% moving data sources, services and storage are distributed across institutions
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{Cosmic Microwave Background}}} \\ \hline	
	      Dynamic resource selection
%dynamic utilization of resources (4)
& Compute on distributed HPC centers \emph{(Data Movement)}
% (sub-set) replication (during application run) \note{What phases of the application do we consider? data acquisition? app run?}\ysnote{Can any of the supercomputers be selected? If so, the data is moved to distributed compute (or it is dynamic resource selection)}
\\
		\hline
\multicolumn{2}{|l|}{{\textbf{Sensor Network Application}}} \\ \hline	
	     Notification of sensor data when in range, data rates and volumes \emph{(Data Movement, Notification)} %data rate, volume dynamic (1-2)
&Aggregating distributed sensor data \emph{(Data Movement, Notification)}  \\
		\hline
\multicolumn{2}{|l|}{{\textbf{Climate (ESGF)}}} \\ \hline	
	      Manage data growth % New data generated over time (ESGF-wide - is this in scope of the application?)
             &
Distributed, replicated catalogs using ESGF software stack (GridFTP, wget/HTTP) \emph{(Data Access,
  Data Discovery, Data Movement, Data Conversion)} \\
  % multiple climate centers (natural), users can request data from ESGF (sub-set replication), coupled
% Can come in via website - cookie / OpenID
% Come in through command line script - grid cert
		\hline
\multicolumn{2}{|l|}{{\textbf{Interactive Exploration of Environmental Data}}} \\ \hline	
On-demand data subsetting, caching and versioning, real-time instrument data \emph{(Data Discovery, Data
  Movement, Data Sources)}
% at moment services are hosted on a single address. There may be farming for load balancing in the future.
% data rates change (1),
% 	    data inputs/outputs change (3), queries change (7)
		&
Distributed Data Services, Access Restrictions \emph{(Data
  Movement, Data Access)}

@@

\\
		\hline
	    % Dynamic Environmental Data on the Web &data operated on changes
	    %     (realtime feeds) (2)
	    %     &natural (sensor sources, simulations), (sub-set) replication\\
	    %     \hline
\multicolumn{2}{|l|}{{\textbf{PowerGrids}}} \\ \hline

@@

		\hline
\multicolumn{2}{|l|}{{\textbf{Fusion (ITER)}}} \\ \hline	
	     Stream and transform data between simulation sites on-demand \emph{(Data Movement, Enrichment)}
%  (multiple distinct data
	    % flows), on-the-fly
	    % transformation (8)
	    &
Run analysis on available, distributed resources \emph{(Data Movement)} %Analysis
% natural, coupled, transient sub-set replicated
            \\
		\hline
\multicolumn{2}{|l|}{{\textbf{Industrial Incident Notification and Response}}} \\ \hline	
	     Acquire different data sources and rates, Processing determined by data
             \emph{(Data Sources, Data Access, Notification, Orchestration)}   %Analysis,  % (1),
	    % data itself changes (2), dynamic data sources (3), processing changes
	    % based on data
            %(5)
&
Distributed data sources, Analysis by agents across institutions \emph{(Data Sources, Data
  Access)} % , Analysis
% natural, coupled
            \\
		\hline
\multicolumn{2}{|l|}{{\textbf{MODIS Data Processing}}} \\ \hline

@@

\multicolumn{2}{|l|}{{\textbf{Distributed Network Intrusion Detection}}} \\ \hline	
Analyze dynamic network traffic rates \emph{(Data Movement)} % , Analysis
	    %  data rate variable (1), data
	    % highly dynamic changing (2), application responds to data (5),
	    % on-the-flight processing (8)
            &
Run analysis at different sites, Distribute generated data \emph{(Data Movement)} % Analysis,
% natural, data/compute tightly coupled
\\
		\hline
%	 \end{longtable}
%\end{footnotesize}

@@

Revisiting the overall issue of infrastructure, which we defined as
the hardware and software needed to run the applications, we
note that the specific hardware isn't a primary issue.  The
applications we've discussed all use a mix of the systems that exist
today, which have aspects of web, grid, and cloud, and can be
considered a hybrid of the three.  The key infrastructure components
are really the software and services on top of the hardware.  In this
section, we've discussed infrastructure to support data management and
data analysis, as well as the tools and systems used to program
applications, and the specific components used for the application in
\S\ref{sec:scenarios}.  However, there is not much commonality in the
software and services that the applications use, so the software and programming infrastructure
we've described here is much more broad than what is used by the
applications we've studied.

@@

\subsection{Observations}
\label{sec:observations}

In this subsection we summarize the source and nature of dynamic data that we observe in
the application surveyed in \S\ref{sec:scenarios}, and the types of
data distribution that we observe in such applications.
